Proofs as Programs in Classical Logic\\Notes

### 2021-09-28 Notes/notes.tex
Added intuitionistic example for simple Curry Howard

\usepackage{amsmath,amssymb,amsthm}
\usepackage{ wasysym }
\usepackage{ stmaryrd }
\usepackage{ mathpartir }
\usepackage{xcolor}

@@

Now to retrieve a value we just need to handle the final exception, i.e.
$$\lambda x_1.\alpha_4\:\mathbf{handle}\: \overline{e}_p(x_2, y_2)(w)\Rightarrow y_2$$
Of course this will not yield an actual value unless $\alpha_1$ yields.\\
We will now consider a minimal inductive example.\pagebreak

@@

\begin{example}\hfill
	\begin{align}
		& p(z, z) & \text{Axiom}\\
		& \neg p(n_2, m_2)\vee p(s(n_2), s(s(m_2)))& \text{Axiom}\\
		& \llb \neg p(n_3, m_3) | x\simeq n_3\rrb & \text{negated Hypothesis}\\
		& \llb \square | x\simeq z\rrb & res.\:(1) (3) [z/n_3, z/m_3]\\
		& \llb \neg p(n_2, m_2) | x\simeq s(n_2)\rrb &res. (2) (3) [s(n_2)/n_3], s(s(m_2))/m_3]\\
		& \text{Activate $\alpha := p(a, b)[a, b]$}&\text{Trigger  (4)}\\
		& \llb n_7\not\prec n_7'\vee p(n_7, m_7) | \mathbf{T}_\alpha^\prec(n_7', m_7)\rrb& \Gamma_{\mathbf{T}_\alpha^\prec}\\
		& \llb n_7\not\prec n_7'| x\simeq s(n_7)\wedge \mathbf{T}_\alpha^\prec(n_7', m_7)\rrb & res. (5) (7) [n_7/n_2, m_7/ m_2] \\
		& \llb n_7\prec s(n_7)|\top\rrb &\Gamma_\prec^s\\
		& \llb \square | x\simeq s(n_7)\wedge\mathbf{T}_\alpha^\prec(s(n_7), m_7)\rrb&res. (8)(9)[s(n_7)/n_7']\\
		& \llb \square | x\simeq n_{10}\rrb &\text{Induction }(4)(9)\\
		& \square &\text{Instantiation }(10)
	\end{align}
\end{example}

\bibliographystyle{acm}
\bibliography{references}

>>>>

\usepackage{amsmath,amssymb,amsthm}
\usepackage{ wasysym }
\usepackage{ stmaryrd }
\usepackage{ mathpartir }
\usepackage{xcolor}
\usepackage{bussproofs}

@@

Now to retrieve a value we just need to handle the final exception, i.e.
$$\lambda x_1.\alpha_4\:\mathbf{handle}\: \overline{e}_p(x_2, y_2)(w)\Rightarrow y_2$$
Of course this will not yield an actual value unless $\alpha_1$ yields.\\
We will now consider a minimal inductive example.

@@

\begin{example}\hfill
	\begin{align}
		& p(z, z) & \text{Axiom}\\
		& \neg p(n_2, m_2)\vee p(s(n_2), s(s(m_2)))& \text{Axiom}\\
		& \llb \neg p(n_3, m_3) | x\simeq n_3\rrb & \text{negated Hypothesis}\\
		& \llb \square | x\simeq z\rrb & res.\:(1) (3) [z/n_3, z/m_3]\\
		& \llb \neg p(n_2, m_2) | x\simeq s(n_2)\rrb &res. (2) (3) [s(n_2)/n_3], s(s(m_2))/m_3]\\
		& \text{Activate $\alpha := p(a, b)[a, b]$}&\text{Trigger  (4)}\\
		& \llb n_7\not\prec n_7'\vee p(n_7, m_7) | \mathbf{T}_\alpha^\prec(n_7', m_7)\rrb& \Gamma_{\mathbf{T}_\alpha^\prec}\\
		& \llb n_7\not\prec n_7'| x\simeq s(n_7)\wedge \mathbf{T}_\alpha^\prec(n_7', m_7)\rrb & res. (5) (7) [n_7/n_2, m_7/ m_2] \\
		& \llb n_7\prec s(n_7)|\top\rrb &\Gamma_\prec^s\\
		& \llb \square | x\simeq s(n_7)\wedge\mathbf{T}_\alpha^\prec(s(n_7), m_7)\rrb&res. (8)(9)[s(n_7)/n_7']\\
		& \llb \square | x\simeq n_{10}\rrb &\text{Induction }(4)(9)\\
		& \square &\text{Instantiation }(10)
	\end{align}
\end{example}
\pagebreak
\subsection{Intuitionistic Example}

For simplicity we will use a single sort (0-level type).\\
Function symbols: $f(-)$\\
Predicate symbols: $P(-), Q(-, -)$\\
Axioms: $\forall x P(x)$, $\forall x\exists y x=f(y)$, $\forall x(P(x)\supset Q(f(x), x))$\\
To Show: $\forall x \exists y Q(x, y)$\\

\noindent Below is a ND-style proof and translation based on the calculus and correspondence presented in~\cite{pfenning2000constructive}.\\

\small{
\AxiomC{}
\RightLabel{\scriptsize Axiom}
\UnaryInfC{$\vdash\forall x\exists y f(y)=x$}
\RightLabel{\scriptsize $\forall E$}
\UnaryInfC{$\vdash\exists y f(y)=x$}
\AxiomC{}
\RightLabel{\scriptsize Axiom}
\UnaryInfC{$\vdash \forall x(P(x)\supset Q(f(x), x))$}
\RightLabel{\scriptsize $\forall E$}
\UnaryInfC{$\vdash P(y)\supset Q(f(y), y)$}
\AxiomC{}
\RightLabel{\scriptsize Axiom}
\UnaryInfC{$\vdash\forall x P(x)$}
\RightLabel{\scriptsize $\forall E$}
\UnaryInfC{$\vdash P(y)$}
\RightLabel{\scriptsize $\supset\hspace{-2pt}E$}
\BinaryInfC{$\vdash Q(f(y), y)$}
\AxiomC{}
\RightLabel{\scriptsize u}
\UnaryInfC{[$\vdash f(y) = x$]}
\RightLabel{\scriptsize subst}
\BinaryInfC{$\vdash Q(x, y)$}
\RightLabel{\scriptsize $\supset\hspace{-2pt}I^u$}
\UnaryInfC{$f(y)=x\vdash Q(x, y)$}
\RightLabel{\scriptsize$\exists I$}
\UnaryInfC{$f(y)=x\vdash \exists y Q(x, y)$}
\RightLabel{\scriptsize $\exists E$}
\BinaryInfC{$\vdash \exists Q(x, y)$}
\RightLabel{\scriptsize $\forall I$}
\UnaryInfC{$\vdash\forall x \exists yQ(x, y)$}

\hspace*{-1cm}
\DisplayProof
}
\vspace{3cm}
\small{
	\AxiomC{}
	\RightLabel{\scriptsize Axiom}
	\UnaryInfC{$\vdash \alpha : \Pi_x\Sigma_y f(y)=x$}
	\RightLabel{\scriptsize $\forall E$}
	\UnaryInfC{$\vdash \alpha(x): \Sigma_y f(y)=x$}
	\AxiomC{}
	\RightLabel{\scriptsize Axiom}
	\UnaryInfC{$\vdash \beta : \Pi_x(P(x)\supset Q(f(x), x))$}
	\RightLabel{\scriptsize $\forall E$}
	\UnaryInfC{$\vdash \beta(y) : P(y)\supset Q(f(y), y)$}
	\AxiomC{}
	\RightLabel{\scriptsize Axiom}
	\UnaryInfC{$\vdash\Gamma\Pi_x P(x)$}
	\RightLabel{\scriptsize $\forall E$}
	\UnaryInfC{$\vdash\gamma(y) : P(y)$}
	\RightLabel{\scriptsize $\supset\hspace{-2pt}E$}
	\BinaryInfC{$\vdash \beta(y)(\gamma(y)) : Q(f(y), y)$}
	\AxiomC{}
	\RightLabel{\scriptsize u}
	\UnaryInfC{[$\vdash\delta : f(y) = x$]}
	\RightLabel{\scriptsize subst}
	\BinaryInfC{$\vdash \beta(y)(\gamma(y))_\delta : Q(x, y)$}
	\RightLabel{\scriptsize $\supset\hspace{-2pt}I^u$}
	\UnaryInfC{$\delta : f(y)=x\vdash\beta(y)(\gamma(y))_\delta : Q(x, y)$}
	\RightLabel{\scriptsize$\exists I$}
	\UnaryInfC{$\delta : f(y)=x\vdash \langle y, \beta(y)(\gamma(y))_\delta\rangle :\Sigma_y Q(x, y)$}
	\RightLabel{\scriptsize $\exists E$}
	\BinaryInfC{$\vdash \textbf{let }\langle y, \delta\rangle = \alpha(x)\textbf{ in }\langle y,\beta(y)(\gamma(y)\delta)\rangle: \Sigma_y Q(x, y)$}
	\RightLabel{\scriptsize $\forall I$}
	\UnaryInfC{$\lambda x.\textbf{let }\langle y, \delta\rangle = \alpha(x)\textbf{ in }\langle y,\beta(y)(\gamma(y))_\delta\rangle:\Pi_x \Sigma_y Q(x, y)$}

	\hspace*{-1cm}
	\DisplayProof
}
\hfill
\\\\\hspace{2cm}We can reduce the last term to $\lambda x.\langle \alpha(x)_1, \beta(\alpha(x)_1)(\gamma(\alpha(x)_1)_{\alpha(x)_2}\rangle$.

\noindent Finally replacing the proof terms by units we obtain $\lambda x.\langle \alpha(x)_1, ()\rangle : \Pi_x\Sigma_y \top$ from which we may further extract $\lambda x. \alpha(x)_1$.

\pagebreak
\bibliographystyle{acm}
\bibliography{references}
